Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-30 PAPER.tex
removed language about Kolopanis exploring a larger parameter space for gamma

We also saw in Section \ref{sec:otherweight} how adding the identity matrix to the empirical covariance can minimize signal loss. We experiment with this idea as well, shown in the bottom panel of Figure \ref{fig:sigloss_modeloop}. The gray and black lines represent power spectrum limits pre and post-signal loss estimation, respectively, as a function of the strength of $\textbf{I}$ that is added to $\widehat{\textbf{C}}$, quantified as the percentage of Tr($\widehat{\textbf{C}})\textbf{I}$ added to $\widehat{\textbf{C}}$. We parameterize this ``regularization strength" parameter as $\gamma$, namely $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + \gamma$Tr$(\widehat{\textbf{C}})\textbf{I}$. From this plot we see that only a small percentage of Tr($\widehat{\textbf{C}})$ is needed to significantly reduce loss. We expect that as the strength of $\textbf{I}$ is increased (going to the left), both the black and gray curves will approach the uniform-weighted case. We also notice that the post-signal loss limit hovers around the uniform-weighted limit for a large range of strengths and while an overall trend from high to low signal loss is seen as the strength increases, there does not appear to be a clear `minimum' that produces the least loss.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (gray) and after (black) signal loss estimation as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). Over $\sim3$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (gray) and after (black) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e. very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (dashed blue) and inverse variance 
weighted case (red); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as solid green. In the revised PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.035 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.035$) as a simple example of regularization that minimizes loss.}
	\label{fig:sigloss_modeloop}
\end{figure*}

In addition to our thermal noise prediction (green) and uniform-weighted power spectrum limit (blue), one additional horizontal line is shown in Figure \ref{fig:sigloss_modeloop} 
in both panels and represents a third regularization technique. This line (red) denotes the power spectrum value, post-signal loss estimation, for inverse variance weighting (multiplying an identity 
matrix element-wise to $\widehat{\textbf{C}}$). This result is single-valued and not a function of the horizontal axis. We see that all three regularization schemes shown (black solid, black dashed, red) perform similarly at 
their best (i.e. when $\sim2$ eigenmodes are down-weighted in the case of the black solid curve). However, for the remainder of this paper, we choose to use the weighting option of $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + 0.035 \,$Tr($\widehat{\textbf{C}})\textbf{I}$, or $\gamma = 0.035$, which we will denote as $\widehat{\textbf{C}}_{\rm eff}$. We choose this weighting scheme merely as a simple example of regularizing PAPER-64 covariances, though we note that in Kolopanis et al. (\textit{submitted}), we choose the most optimal regularization parameter for our final limits for every redshift bin independently. We also note that all of the regularizations described here perform similarly to the uniformly weighted case. However, we find that the regularization of empirical covariances does outperform the uniform weighted case for redshift bins in which there are stronger foregrounds, as presented in Kolopanis et al. (\textit{submitted}).

A revised PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{\rm eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, the dashed black curve represents our upper limit on EoR, post-signal loss estimation. Black and gray points correspond to positive and negative power spectrum values respectively (pre-signal loss estimation), with 
$2\sigma$ errors bars. Also plotted are the uniform-weighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). In Kolopanis et al. (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$ Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.pdf}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{\rm eff}$. The weighted power spectrum, pre-signal loss estimation, is shown by the black and gray points (corresponding to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. There is a small amount of loss using this regularization. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. In Kolopanis et al. (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$ Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.}
	\label{fig:ps1_data}
\end{figure*}

>>>>

We also saw in Section \ref{sec:otherweight} how adding the identity matrix to the empirical covariance can minimize signal loss. We experiment with this idea as well, shown in the bottom panel of Figure \ref{fig:sigloss_modeloop}. The gray and black lines represent power spectrum limits pre and post-signal loss estimation, respectively, as a function of the strength of $\textbf{I}$ that is added to $\widehat{\textbf{C}}$, quantified as the percentage of Tr($\widehat{\textbf{C}})\textbf{I}$ added to $\widehat{\textbf{C}}$. We parameterize this ``regularization strength" parameter as $\gamma$, namely $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + \gamma$Tr$(\widehat{\textbf{C}})\textbf{I}$. From this plot we see that only a small percentage of Tr($\widehat{\textbf{C}})$ is needed to significantly reduce loss. We expect that as the strength of $\textbf{I}$ is increased (going to the left), both the black and gray curves will approach the uniform-weighted case. We also notice that the post-signal loss limit hovers around the uniform-weighted limit for a large range of regularization strengths and while an overall trend from high-to-low signal loss is seen as the strength increases, there does not appear to be a clear `minimum' that produces the least loss.

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$\,$h$ Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (gray) and after (black) signal loss estimation as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). Over $\sim3$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (gray) and after (black) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $\gamma = 10^{-4}$ on the right (i.e. very little regularization) to $\gamma = 1$ on the left (see main text for the definition of $\gamma$). Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (dashed blue) and inverse variance 
weighted case (red); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as solid green. In the revised PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}}_{\rm eff} \equiv 0.035 \, $Tr($\widehat{\textbf{C}})\textbf{I} + \widehat{\textbf{C}}$ ($\gamma = 0.035$) as a simple example of regularization that minimizes loss, and note that the power spectrum limits are roughly constant across a large range of values of $\gamma$.}
	\label{fig:sigloss_modeloop}
\end{figure*}

In addition to our thermal noise prediction (green) and uniform-weighted power spectrum limit (blue), one additional horizontal line is shown in Figure \ref{fig:sigloss_modeloop} 
in both panels and represents a third regularization technique. This line (red) denotes the power spectrum value, post-signal loss estimation, for inverse variance weighting (multiplying an identity 
matrix element-wise to $\widehat{\textbf{C}}$). This result is single-valued and not a function of the horizontal axis. We see that all three regularization schemes shown (black solid, black dashed, red) perform similarly at 
their best (i.e. when $\sim2$ eigenmodes are down-weighted in the case of the black solid curve). However, for the remainder of this paper, we choose to use the weighting option of $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} + 0.035 \,$Tr($\widehat{\textbf{C}})\textbf{I}$, or $\gamma = 0.035$, which we will denote as $\widehat{\textbf{C}}_{\rm eff}$. We choose this weighting scheme merely as a simple example of regularizing PAPER-64 covariances, noting that the power spectrum upper limit remains roughly constant for a broad range of values of $\gamma$. We also note that all of the regularizations described here perform similarly to the uniformly weighted case. However, we find that the regularization of empirical covariances does outperform the uniform weighted case for redshift bins in which there are stronger foregrounds, as presented in Kolopanis et al. (\textit{submitted}).

A revised PAPER-64 power spectrum (using only one baseline separation type and $\widehat{\textbf{C}}_{\rm eff}$) is shown in Figure 
\ref{fig:ps1_data}. Again, the dashed black curve represents our upper limit on EoR, post-signal loss estimation. Black and gray points correspond to positive and negative power spectrum values respectively (pre-signal loss estimation), with 
$2\sigma$ errors bars. Also plotted are the uniform-weighted power spectrum upper limit (dashed blue) and theoretical prediction of 
noise (solid green). In Kolopanis et al. (\textit{submitted}), multiple baselines are used in the analysis to produce a $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$ Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.pdf}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{\rm eff}$. The weighted power spectrum, pre-signal loss estimation, is shown by the black and gray points (corresponding to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. There is a small amount of loss using this regularization. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. In Kolopanis et al. (\textit{submitted}), multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$ Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.}
	\label{fig:ps1_data}
\end{figure*}
